---
title: "Ames Housing Project - Exploratory Analysis"
author: "Quyen Duong, Amin Jamali, Camille Nigon, Pascal Waser"
date: "Last edited `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false 
      smooth_scroll: false
    toc_depth: 4
    number_section: true 
---
<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
}
code.r{
  font-size: 11px;
}
pre {
  font-size: 11px
}
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 32px;
  font-weight: bold;
  text-align: center;
  color: #2A5A7A;
  opacity: 0.8;
}
h4.author { 
  font-size: 15px;
  color: #468faf;
  text-align: center;
}
h4.date { 
  font-size: 16px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h1 { /* Header 1 */
  font-size: 22px;
  color: Black;
  font-weight: bold;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: Black;
  font-weight: bold
}
h3 { /* Header 3 */
  font-size: 15px;
  color: Black;
  font-weight: bold
}

</style>

```{r setup, include=FALSE}
# Set the default mode for all the chunks
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.align = "center", 
                      fig.height = 5, fig.width = 8)
```

```{r load-packages, include = FALSE}
# Load packages
library(dplyr)
library(ggplot2)
library(tidyverse)
library(scales)                      # formatting numbers on axis ggplot
library(gridExtra)
library(tidyr)
library(readr)
library(knitr)
library(lubridate)                   # to format date
library(egg)                         # plot layout 
library(PerformanceAnalytics)        # correlation matrix
library(GGally)                      # correlation matrix
library(corrplot)                    # correlation matrix
library(Hmisc)                       # correlation matrix
library(psych)                       # summary statistics
library(caret)                       # split data train-test
library(broom)                       # visualize fitted linear with categorical variable
library(MASS)
library(conflicted)                  # set conflict
library(ggthemes)
library(grid)
library(leaps)                       # stepwise selection

# Set conflict preference
conflict_prefer("filter", "dplyr")
conflict_prefer("select", "dplyr")

# Deactivate scientific notation
# options(scipen = 999)
```

```{r set-up-theme-for-ggplot, include = FALSE}
my_theme_general <- theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
        text = element_text(size = 10, color = "#495057"),
        panel.grid.major = element_line(linetype = "dashed"),
        panel.grid.minor = element_blank()
    ) 

my_theme_bar_chart <- theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5),
        text = element_text(size = 10, color = "#495057"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(linetype = "dashed"),
        panel.grid.major.y = element_blank()
    )

my_theme_heat_map <-  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    text = element_text(size = 10, color = "#495057"),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
    )

my_blank_theme <- theme_minimal()+
  theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.border = element_blank(),
  panel.grid=element_blank(),
  axis.ticks = element_blank(),
  plot.title=element_text(size=14, face="bold")
  )
```

# Motivation and purpose

The study attempts to predict the house price in Ames, Iowa. To predict the house price, several machine learning methods will be used such as linear model (LM), Generalised Linear Model (GLM), Generalised Additive Model (GAM), Neural Network (NN), Support Vector Machine (SVM) and Approximate Bayesian Computation. Yet, before predicting price and applying models, it's vital to explore and understand the data. 

# Data comprehension

```{r load-data, include = FALSE}
# Load data
d_house <- read.csv("dataset/AmesHousing.csv", stringsAsFactors = TRUE) %>% 
  as_tibble()

# Check the variables
glimpse(d_house)

# Write a new Excel with all variable name and check if they are important to the house price
# names(d_house) %>% 
#   as_tibble() %>% 
#   writexl::write_xlsx("variables_intuition_analysis.xlsx")

# Exclude the irrelevant variables
d_house <- d_house %>% 
  select(-c(Order, PID))
```

There are 82 variables and 2930 observations in the data set. As there are many variables, we first have to examine and make a intuition analysis on every variable and its relevance to the main variable ´SalePrice´. Although this process can be very time-consuming, it shed the light to grasp the data better. How did we do that? Below are the step explanation:

* Reading and understanding every variable description
* Make a new Excel with two columns: variable name and importance to sale price (Yes/No). The decision of Yes/No is grounded on intuition, expectation and some house-price related articles. To be more specific, we question these following questions: which factor matters when we buy a house? How important would this factor be? Does this factor appear in other variables?

The intuition-based Excel file can be found in the ´variables_intuition_analysis.xlsx´. At the first glimpse, we marked these might-be-relevant features to the SalePrice: MS.Zoning, Lot.Area, Street, Lot.Shape, Neighborhood, Bldg.Type, House.Style, Overall.Qual, Year.Built, Foundation, Total.Bsmt.SF, TotRms.AbvGrd, Mo.Sold, Yr.Sold, Garage.Cars, Sale.Condition.

After thinking of potential features, let's inspect the response variable in the next part. 

# The response variable - Sale price 

The scope of this analysis is to predict the housing sales price based on appropriate features. First, we plot the histogram and box plot of the sale price (i.e. the dependent variable).

```{r histogram-sales-price}
# Histogram
histogram_sale_price <- ggplot(d_house, mapping = aes(x = SalePrice)) +
  geom_histogram(aes(y=..density..), color="black", fill = "grey") +
  geom_density(alpha=.2, color ="blue") +
  scale_x_continuous(labels = comma) +
  labs(title = "Histogram of sale price") +
  my_theme_general

# Boxplot
boxplot_sale_price <- ggplot(d_house, mapping = aes(x = "", y = SalePrice)) +
  geom_boxplot(color = "black", fill = "grey") +
  scale_y_continuous(labels = comma) +
  coord_flip() +
   xlab("") +
  my_theme_general

# Display two plots together
egg::ggarrange(histogram_sale_price, boxplot_sale_price, heights = 2:1)

# Descriptive statistics summary of SalePrice
stats_SalePrice <- psych::describe(d_house$SalePrice)
```

The distribution of sale prices displays right-skewed with a long tail. The skew of `r stats_SalePrice$skew` confirms that the data are highly confirmed. Besides, the Kurtosis is `r stats_SalePrice$kurtosis`, i.e. Leptokurtic with a longer distribution and fatter tails. Some houses are very expensive. We have to carefully consider whether these expensive houses are outliers. Instead of dropping outliers, we apply the log transformation for sale prices. 

From the beginning to this point, we just follow the intuition for variable selection. However, this method can be too subjective. Hence, a more objective and systematic way need to be executed as the section below. Nevertheless, we still consider our initial thoughts as a double-check criteria and critical thinking in supplement with the multivariate study.    

# Numerical variable analysis

## Correlation matrix 

```{r corr-matrix}
# Overview correlation matrix of all numerical variables
d_house_numeric <- d_house %>% 
  select(where(is.numeric))

# Missing value check in numerical features -> not so many NA
round(colMeans(is.na(d_house_numeric)), 2) %>%  
  as.data.frame()

# Overview histograms of all numerical variables
# hist.data.frame(d_house_numeric)

# Create correlation matrix, use = complete to obmit NA, because not so many NA
m_cor <- cor(d_house_numeric, use = "complete") 

# Correlogram
corrplot(m_cor, method = "color", addCoef.col="black",
         order = "alphabet", number.cex=0.35, tl.cex = 0.55)
```

The correlation matrix shows some dark blue areas where variables are highly correlated to each other. To be more specific, `Garage.Area` is highly correlated `Garage.Cars` or other pairs such  `Gr.Liv.Area` with `TotRms.AbvGrd`; `Garage.Yr.Blt` with `Year.Built`. These pairs indicate the strong correlation and multi-collinearity. When we check these variables' meaning, they give almost similar information. Here, the correlation matrix is a crucial method to detect the multi-collinear problems in the feature selection process.

```{r corr-SalePrice}
# Calculate correlation matrix for all numerical features against SalePrice
m_cor_SalePrice <- round(
  cor(d_house_numeric, d_house["SalePrice"], use = "complete", ) ,2)

# Transform into tibble and sort correlation in a descending order
d_cor_SalePrice <- tibble(var_name = row.names(m_cor_SalePrice), 
       SalePrice = m_cor_SalePrice[, "SalePrice"]) %>% 
  arrange(desc(SalePrice))
d_cor_SalePrice 
```

Both variables have long tails to the right and many outliers. 

Remarkably, `Lot.Area`, the lot size in square feet is one of the variable we considered as important at the first sight but its correlation to the `SalePrice` is only 0.31. Let's look at the scatter plot of these two variables:

```{r scatter-SalePrice-Lot.Area}
# Scatter plot SalePrice & Lot.Area
scatter_price_lot <- ggplot(d_house, aes(x = Lot.Area, y = SalePrice)) + 
  geom_point() 

# Scatter plot SalePrice & Lot.Area but change the x & y range
scatter_price_lot2 <- ggplot(d_house, aes(x = Lot.Area, y = SalePrice)) + 
  geom_point() +
  xlim(0, 50000) +
  ylim(0, 500000)

# Plot the two side-by-side
egg::ggarrange(scatter_price_lot, scatter_price_lot2, ncol=2, nrow = 1)
```

The two plots show the same information but the right scatter displays the removal of outliers. The `Lot.Area` data points stay mostly at the smaller value and their price varies vertically for the same lot space.

## Top 10 variables correlated to SalePrice

The top 10 numerical features with the absolute highest correlation with the `SalePrice` are now selected and we produce a SalePrice correlation matrix as below:  

```{r top10-numerical-correlation}
# Make df of top 10 numerical variables with the absolute highest correlation to SalePrice
d_cor_top_10 <- d_cor_SalePrice %>%  
  mutate(abs_corr = abs(SalePrice)) %>% 
  arrange(desc(abs_corr)) %>% 
  slice(1:10)

# Make vector with variable name
v_top_10_cor <- d_cor_top_10 %>% 
  # Select the variable names and convert to vector 
  pull(var_name)

# Correlation matrix of top 10
m_cor_top_10 <- cor(d_house_numeric %>% 
               select(all_of(v_top_10_cor)), 
             use = "complete") 

# Correlogram
corrplot(m_cor_top_10, method = "color", addCoef.col="black", type = 'lower',
         order = "alphabet", number.cex=0.55, tl.cex = 0.55)

# Quick look on the scatter plot in pair for the top 10
ggpairs(d_house_numeric %>% select(all_of(v_top_10_cor)))
```

From the new correlation matrix, some important findings are found:

* `Garage.Area` and `Garage.Cars` are correlated highly with the dependent variable but they have the same information. Here, we keep the `Garage.Cars` as its higher correlation.  
* `SalePrice` is strongly correlated to `Gr.Liv.Area`, `Overall.Qual`, `Total.Bsmt.SF` 
* `Full.Bath` seems to be strange here
* `BsmtFin.SF.1` and `Total.Bsmt.SF` display the related information, we choose `Total.Bsmt.SF` in this case
* `TotRms.AbvGrd` and `Gr.Liv.Area` are similar. Here, we choose `Gr.Liv.Area`
* `Year.Built`: 0.55 - moderately correlated to `SalePrice` 

After choosing the variables, let's make scatter plots to check the relationship between SalePrice and other variables

```{r scatter-plots-selected-var-vs-SalePrice}
# Selected numerical variables
d_select_num <- d_house_numeric %>% 
  select(SalePrice, Garage.Cars, Gr.Liv.Area, Overall.Qual, Total.Bsmt.SF, Full.Bath, Year.Built)

# list of selected variables 
l_vars <- c("Garage.Cars",  "Gr.Liv.Area", "Overall.Qual", "Total.Bsmt.SF", "Full.Bath",  "Year.Built")

# empty list for scatter plot
plotlist <- list()

# For loop to get several scatter plots at once
for (idx in 1:length(l_vars)) {
  this_var <- l_vars[idx]
  p <- ggplot(d_select_num, aes_string(x = this_var, y = "SalePrice")) +
    geom_point() +
    geom_smooth(method = lm)
  plotlist[[idx]] <- p
} 

# 6 scatter plots at once
egg::ggarrange(plotlist[[1]], plotlist[[2]], 
               plotlist[[3]], plotlist[[4]],
               plotlist[[4]], plotlist[[6]], ncol=3, nrow = 2)
```

In general, the plots indicate that all of these variables have a linear positive relationship with the dependent variable `SalePrice`. 

# Missing value

It's important to inspect the missing values in each variable. How is the ratio of missing value in each feature? What does the missing value mean in each case? Are missing values random? These questions are vital when we deal with the missing value. Removing them immediately without reflecting can cause data size reduction and bias.

```{r missing-values}
# Number of NA values in each column
d_number_missing_values <- colSums(is.na(d_house)) %>%  
  as.data.frame() 
colnames(d_number_missing_values) <- c("number")

# Percentage of NA values in each column
d_perc_missing_values <- round(colMeans(is.na(d_house)), 2) %>%  
  as.data.frame()
colnames(d_perc_missing_values) <- c("ratio")
d_perc_missing_values

# A summary table of missing values in each column
d_summary_missing_value <- cbind(d_number_missing_values, d_perc_missing_values)

# Change the summary missing values to tibble
d_summary_missing_value <- tibble(var_name = row.names(d_summary_missing_value), 
       n_missing = d_summary_missing_value[, "number"],
       perc_missing = d_summary_missing_value[, "ratio"]) %>% 
  arrange(desc(perc_missing)) 
d_summary_missing_value

# Exclude the columns with 49% missing value
d_house <- d_house %>% 
  select(-c(Alley, Fireplace.Qu, Pool.QC, Fence, Misc.Feature))
```

After examine the missing ratio, here are some decisions:

* Excluding all the columns with more than 49% of missing values, namely `Pool.QC`, `Misc.Feature`, `Alley`, `Fence` and `Fireplace.Qu`.Filling NA value is not applied because these variables are not so important and not the main factors to think about when we buy a house.
* Group of variables with "Garage": 5% are missing in these columns. Since we chose only `Garage.Cars` for the model in the previous part, we do not care about the rest.
* Group of variables with "Bsmt": similar principle as the group of "Garage"

# Categorical variables

In this section, we check the effects of different categorical features that are potentially relevant for the model. Before applying function drop1(), we visualize SalePrice with several categorical variables

## Visualizations of categorical variables

```{r Boxplot-Fuction}
# Make a function to draw all box plots
fill <- "#FFFFFF"
line <- "#000000"

boxplot_theme <- function(variable_input, title, variable) {
  output <- ggplot(d_house, aes(x = variable_input, y = SalePrice)) +
    geom_boxplot(fill = fill, colour = line) +
    ggtitle(title) +
    theme_minimal() +
    scale_y_continuous(labels = comma) +
    xlab(paste0("\n", variable)) +
    ylab("Price\n") +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1),
          legend.position = "bottom", legend.direction = "horizontal",
          legend.box = "horizontal",
          legend.key.size = unit(1, "cm"),
          plot.title = element_text(family="Tahoma", hjust = 0.5),
          text = element_text(family = "Tahoma"),
          axis.title = element_text(size = 10),
          legend.text = element_text(size = 9),
          legend.title=element_text(face = "bold"))

  return(output)
}  
```

### MS.Zoning

`SalePrice` seems to vary among different `MSZoning`. Here, `MSZoning` identifies the general zoning classification of the sales such as agriculture, commercial, industrial. Especially Residential with low density (RL) and residential with medium density are strongly right skewed.

```{r ms.zoning}
# Boxplot price in ms zonings
title <- "General Zoning Classification"
variable <- "MS Zoning"
boxplot_theme(d_house$MS.Zoning, title, variable)
```

### Lot Shape

The general shapes of the property appear to be similar. With the exception that properties with a regular shape has a lower value compared to an irregular shape.

```{r LotShape}
# Boxplot price in different lot shapes
title <- "General shape of property"
variable <- "Lot Shape"
boxplot_theme(d_house$Lot.Shape, title, variable)
```

### Neighborhood

We identified especially in the neighborhoods strong variances regarding the prices. Especially Northridge (NoRidge) has a high price per property and shows some outliers in the direction of an higher price. 

```{r Neighborhood}
# Boxplot price in different neighborhoods
title <- "Physical locations within Ames city limits"
variable <- "Neighborhood"
boxplot_theme(d_house$Neighborhood, title, variable)
```

### Bldg.Type

The type of dwelling or house shows no radical differences for each category. However, Single-family Detached (1Fam) has the highest potential for an higher price.

```{r Bldg.Type}
# Boxplot price in different building types
title <- "Type of building"
variable <- "Building Type"
boxplot_theme(d_house$Bldg.Type, title, variable)
```

### House.Style

The same for House Style. The sales price for one story or two story is not strong impacted by the house style. 

```{r House.Style}
# Boxplot price in different house styles
title <- "Style of dwelling"
variable <- "HouseStyle"
boxplot_theme(d_house$House.Style, title, variable)
```

### Foundation

The `Foundation` variable consists of Brick & Tile, Cinder Block, Poured Contrete, Slab, Stone, Wood. The price seems to differ in different type of foundation but again, this can be a factor that was influenced by the `Year.Built`.

```{r Foundation}
# Boxplot price in different house styles
title <- "Foundation"
variable <- "Foundation"
boxplot_theme(d_house$Foundation, title, variable)
```

### Sale conditions

This variable indicate the condition of sale such as normal sale, abnormal sale (trade, short sale), sale between family members, adjoing land purchases etc. The median of the sales price is the highest among the Sales condition which was not completed when last assessed (associated with new homes). 

```{r Sale conditions}
title <- "Condition of sale"
variable <- "Sale Condition"
boxplot_theme(d_house$Sale.Condition, title, variable)
```

## Testing the effect of categorical variables

After several visualizations and descriptive analysis, let's fit the model with all highly potential categorical variables and then apply drop1() function to test their significance.

```{r drop1-categorical}
# Select all categorical might be relevant
d_house_selected_categorical <- d_house %>% 
  select(MS.Zoning, Neighborhood, Street, House.Style, Sale.Condition)

# Fit model with only these categorical variables
lm_cate_1 <- lm(log(SalePrice) ~ MS.Zoning + Neighborhood + Street + Foundation 
                + House.Style + Sale.Condition, data = d_house)

# Check their significance effects
drop1(lm_cate_1, test = "F")
```

From the p-value obtained, `Street` does not seem to have a relevant effect on the response variable.

## Variable selection

As we have seen so far, the linear model describes relationship between the response variable $Y$ and the predictors $X_1,X_2,...X_n$. However, not all explanatory variables play an important role in predicting the random variable. Now, the question is if omitting a predictor would affect the degree to which the model fits the data, substantially deteriorated or not? 

### Stepwise selection

Command `regsubsets` from the library `leaps` runs the whole procedure automatically. Wherever TRUE is written, the corresponding explanatory variable is contained in model.

```{r step-wise-selection}
# Choose all variables exclude some variables starting with "Bsmt" because cannot run stepwise selection
d_exclude <- d_house %>% 
  select(!starts_with("Bsmt"))

# Fit all variables
lm_full <- lm(log(SalePrice) ~ ., data = d_exclude)

# Forward selection
reg_model <- regsubsets(log(SalePrice) ~ ., data = d_exclude, na.action=na.exclude,
                  method = "forward", nvmax = 11)
reg_sum <- summary(reg_model)

reg_sum$which
```

Based on step wise selection, these are the features containing in the model: Gr.Liv.Area, Overall.Qual, Overall.Cond, Year.Built

### Testing the effects all potential features together

The function `drop1()` gives an overview about the comparison of original model to the model where one explanatory variable was omitted.  

```{r lm-fit-all-potential}
# Fitting all potential features together
lm_fit_select <- lm(log(SalePrice) ~ Gr.Liv.Area + MS.Zoning + Neighborhood + 
                      Overall.Qual + Year.Built + Foundation + Total.Bsmt.SF  + 
                      Garage.Cars + Sale.Condition, data = d_house)

# Testing all potential features together
drop1(lm_fit_select, test = "F")

# Summary of the model 
summary(lm_fit_select)
```

In this case, the `Neighborhood` feature has 28 levels and many of them are not significant. For the location topic, since we take `Ms.Zoning`, we will exclude the `Neighborhood`. 

